---
title: "BLM_popbio_MPM"
author: "Michelle DePrenger-Levin"
date: "6/3/2021"
output: html_document
---

load packages
```{r}
rm(list=ls())

library(popbio)
library(ggplot2)
library(MuMIn)
library(lme4)
library(patchwork)

```

Gather demographic data 
```{r}
# BLM_erbr_popbio <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_BLM_demography_20210603.csv")

BLM_erbr_newplts <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_BLM_newplts.csv")
#   Year.num TransectNew.num num.newPlts num.newPlts1 Year         SiteName  Site
# 1        1              11          NA            2 2016 CastleGarden_BLM  CG_1
# 2        1              12          NA            0 2016 CastleGarden_BLM CG_10


# removed rows before alive but kept first dead, was BLM_erbr.2 in "2021_integratedPopulationModels_BLM-DBG-CNHP_formatdata.R" or dats_BLM need an NA for stage and new for Fate for new
BLM_erbr_lag <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_BLM_demolag_20210603.csv")
#    X Year         SiteName Transect Meter Tag.. Stage Site Data_id        Tag save surv lagsrtstg lagforsurv TransectNew InflYesNo
# 1  1 2016 CastleGarden_BLM        1     0  110*     1 CG_1       1  CG_1_110*    1    1         0          0        CG_1         1
# 2  4 2016 CastleGarden_BLM        3     8  2300     1 CG_3     451  CG_3_2300    1    1         0          0        CG_3         1
#   TransectNew.num Year.num meanTmean.summer TotPpt.summer meanTmean.fall TotPpt.fall meanTmean.winter TotPpt.winter
# 1              11        1          13.9795       107.035        11.6825       71.93           -0.676         63.75
# 2              14        1          13.9795       107.035        11.6825       71.93           -0.676         63.75
newPlts2 <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_newPlts2.csv")

transLevels <- unique(BLM_erbr_lag$Site[with(BLM_erbr_lag, order(SiteName, Transect))])

# by numbered transects 1:87 and years 1:5
# newPlts <- read.csv( "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_BLMnewplts_climate_20210601.csv")
         
# climate_erbr <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_BLM_climate_20210601.csv")

# seasonAllsites <- read.csv(file = "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/2021_Eriogonum-brandegeei_climate.csv")
# seasonAllsites <- seasonAllsites[,-1]

## USFWS climate variables WaterDeficit over different months
CD <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2021_Eriogonum-brandegeei_SSA/2021_WaterDeficitData_Eriogonum-brandegeei_AlexKasdin-USFWS.csv")
head(CD)

# Compare to DBG
date <- "2021-08-27"
name <- as.character("P1k")
DBG_GP <- readRDS(file=paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/2021_Eriogonum-brandegeei_integratedpopulationmodels/erbrLams_noYRE", name, date, sep="_"))

WD <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2021_Eriogonum-brandegeei_SSA/2021_WaterDeficitData_Eriogonum-brandegeei_AlexKasdin-USFWS_reformat.csv")

# Observed values
WD2 <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2021_Eriogonum-brandegeei_SSA/ObservedWaterDeficit_Eriogonum-brandegeei_1979_2020.csv")

head(WD[WD$Months == "Annual",]);head(WD2[WD2$Year > 2002,])

# reshape(dats_BLM, idvar = names(dats_BLM)[c(1:8,12:15)], timevar = "Group.2", direction = "wide") # 

# climate_wide <- reshape(seasonAllsites, idvar = c("Group.3","Group.1"), timevar = "Group.2", direction = "wide")
# names(climate_wide) <- c("Year","Site","Tmean.summer","Ppt.summer","Tmean.winter","Ppt.winter","Tmean.fall","Ppt.fall")
# 
# table(climate_wide$Site)


# names(climate_wide)[3:8]
# climate_wide_scaled <- climate_wide
# climate_wide_scaled[names(climate_wide)[3:8]] <- lapply(climate_wide[names(climate_wide)[3:8]], scale) # dividing centered (subtracted means)
# head(climate_wide_scaled[climate_wide_scaled$Year > 2015,])

# some first year deads got through !!fixed, duplicate tags!! (from 2021_IntegratedPopulationModels_BLM-DBG-CNHP_formatdata)
# BLM_erbr_lag <- BLM_erbr_lag[!(BLM_erbr_lag$Year == 2016 & is.na(BLM_erbr_lag$Stage)),]

# BLM_erbr_lag$SiteName <- "BigBend_BLM"
# BLM_erbr_lag$SiteName[grep("CG", BLM_erbr_lag$Site)] <- "CastleGarden_BLM"
# BLM_erbr_lag$SiteName[grep("DGBLM", BLM_erbr_lag$Site)] <- "Droney_BLM"
# BLM_erbr_lag$SiteName[grep("GPQ", BLM_erbr_lag$Site)] <- "GardenParkQuarry_BLM"
# 
# BLM_erbr_lag_scaled <- merge(BLM_erbr_lag[,-1], climate_wide_scaled, 
#                              by.x = c("SiteName","Year"),
#                              by.y = c("Site","Year"))



BLM_erbr_lag_scaled <- BLM_erbr_lag
BLM_erbr_lag_scaled$StageName <- "dead"
BLM_erbr_lag_scaled$StageName[BLM_erbr_lag_scaled$Stage == 0] <- "vegetative"
BLM_erbr_lag_scaled$StageName[BLM_erbr_lag_scaled$Stage == 1] <- "reproductive"
BLM_erbr_lag_scaled$StageName[BLM_erbr_lag_scaled$Data_id %in% newPlts2$Data_id] <- "seedling" ## all vegetative, good!

## Checks
# BLM_erbr_lag_scaled[BLM_erbr_lag_scaled$Tag == newPlts2$Tag[1],] # new in 2019 in BB, that's good
# BLM_erbr_lag_scaled[BLM_erbr_lag_scaled$Tag == newPlts2$Tag[2],] # new in 2020 BB

table(BLM_erbr_lag_scaled$SiteName,BLM_erbr_lag_scaled$Year, BLM_erbr_lag_scaled$StageName)


## Now there's no Stage for assigning StageName. And this shouldn't be anyway, is summarized over year and transect, right? 
# BLM_erbr_newplts$StageName <- "dead"
# BLM_erbr_newplts$StageName[BLM_erbr_newplts$Stage == 0] <- "vegetative"
# BLM_erbr_newplts$StageName[BLM_erbr_newplts$Stage == 1] <- "reproductive"
# 
# 
# table(BLM_erbr_newplts$SiteName,BLM_erbr_newplts$Year, BLM_erbr_newplts$StageName)

```

Format for popbio projection.matrix  
Use the lag from BLM_erbr_lag to determine the previous time with data to base the missing stage off of
```{r}

BLM_erbr_lag_scaled <- BLM_erbr_lag_scaled[,c("SiteName","Site","Year","Tag","Stage","StageName", "surv", "lagsrtstg", "lagforsurv",
                                              ## NEED lagsrtstg for imputing!
                                              "meanTmean.summer","meanTmean.fall","meanTmean.winter",
                                              "TotPpt.summer","TotPpt.fall","TotPpt.winter")]


head(BLM_erbr_lag_scaled)
BLM_erbr_lag_scaled[,c("meanTmean.summer", "meanTmean.fall","meanTmean.winter", 
                       "TotPpt.summer", "TotPpt.fall", "TotPpt.winter")] <- scale(BLM_erbr_lag_scaled[,c("meanTmean.summer", "meanTmean.fall","meanTmean.winter", 
                       "TotPpt.summer", "TotPpt.fall", "TotPpt.winter")])


# previousYrseedling <- BLM_erbr_lag_scaled[BLM_erbr_lag_scaled$StageName == "seedling",c(1:6)]
# previousYrseedling$StageName <- NA
# previousYrseedling$Year <- previousYrseedling$Year - 1
# previousYrseedling$Stage <- NA
# ## Add in climate from previous year
# previousYrseedling <- merge(previousYrseedling, BLM_erbr_lag_scaled[,c("SiteName","Site","Year","Tag")], 
#                             by = c("SiteName","Site","Year","Tag"))
# # previousYrseedling[,c(7:15)] <- NA
# head(previousYrseedling)
# table(previousYrseedling$Year)
# 
# BLM_erbr_lag_scaled <- rbind(BLM_erbr_lag_scaled, previousYrseedling)
tail(BLM_erbr_lag_scaled)
BLM_erbr_lag_scaled[BLM_erbr_lag_scaled$StageName == "seedling",]

## Merge all.y will maybe keep all the seedlings and add NA to year before? 
BLM_erbr_scaled <- subset(merge(BLM_erbr_lag_scaled, BLM_erbr_lag_scaled, by = c("SiteName","Site","Tag")), 
                          sort=FALSE, Year.x == Year.y - 1)
head(BLM_erbr_scaled)
BLM_erbr_scaled[is.na(BLM_erbr_scaled$SiteName),]
BLM_erbr_scaled[BLM_erbr_scaled$StageName.y == "seedling",] ## Need to remove. If not seen in 2018 at BB 
BLM_erbr_scaled <- BLM_erbr_scaled[BLM_erbr_scaled$StageName.y != "seedling",]

BLM_erbr_scaled[BLM_erbr_scaled$StageName.x == "seedling",] # I think just an artifcat, not real NAs
table(BLM_erbr_scaled$StageName.y,BLM_erbr_scaled$StageName.x)  ### STILL only 1 .y == "seedling"


```


# just use the models from DBG data 
Probability of reproducing: 
     i.e. stage_t ~ stage_t-1 (size for DBG) + winter precip_t + winter temp_t + summer precip_t
          stage.y ~ stage.x                  + Ppt.winter.y      + Tmean.winter.y + Ppt.summer.y             
```{r}
# BLM_erbr_scaled has .x and next year .y
# BLM_erbr_lag_scaled is just a row per year and tag

ggplot(BLM_erbr_scaled, aes(meanTmean.winter.y, TotPpt.summer.y))+
  geom_point()+
  stat_smooth(method = "lm")

ggplot(BLM_erbr_scaled, aes(meanTmean.winter.y, TotPpt.winter.y))+
  geom_point()+
  stat_smooth(method = "lm")


# modelimpute <- glmer(Stage.y ~ Stage.x + Ppt.winter.y + Tmean.winter.x + Ppt.summer.y +
modelimpute <- glmer(Stage.y ~ Stage.x + TotPpt.winter.y + meanTmean.winter.y + TotPpt.summer.y +
                       (1 | Year.y) + (1 | SiteName), 
                     # data = BLM_erbr_scaled_naomit, # BLM_erbr_naomit, 
                     data = BLM_erbr_scaled[complete.cases(BLM_erbr_scaled),],
                     family = binomial(link = "logit"), 
                     na.action = na.fail)

summary(modelimpute)
```

```{r}
# The precip and temperature a full year before 
TempWinter <- ggplot(BLM_erbr_scaled[complete.cases(BLM_erbr_scaled),], # BLM_erbr_scaled_naomit, 
                     aes(meanTmean.winter.y, Stage.y, colour = as.factor(Stage.x)))+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)+
  theme_bw()+
  scale_colour_discrete(name = expression(Stage[t-1]),
                        labels = c("Vegetative","Reproductive"))+
  ylab(expression(Stage[t]))+
  xlab(expression("Mean winter temperature"[t]))


# More likely to be reproductive as Ppt.summer increases when already reproductive previous census
PptSummer <- ggplot(BLM_erbr_scaled[complete.cases(BLM_erbr_scaled),], # BLM_erbr_scaled_naomit, 
                    aes(TotPpt.summer.y, Stage.y, colour = as.factor(Stage.x)))+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)+
  theme_bw()+
  scale_colour_discrete(name = expression(Stage[t-1]),
                        labels = c("Vegetative","Reproductive"))+
  ylab(expression(Stage[t]))+
  xlab(expression("Summer Precipitation"[t]))

PptWinter <- ggplot(BLM_erbr_scaled[complete.cases(BLM_erbr_scaled),], # BLM_erbr_scaled_naomit,
                    aes(TotPpt.winter.y, Stage.y, colour = as.factor(Stage.x)))+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)+
  theme_bw()+
  scale_colour_discrete(name = expression(Stage[t-1]),
                        labels = c("Vegetative","Reproductive"))+
  ylab(expression(Stage[t]))+
  xlab(expression("Winter Precipitation"[t]))


ggsave("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/2021_Eriogonum-brandegeei_integrated_figures/ImputeBLMdata.jpg",
       TempWinter/PptSummer/PptWinter,
width=120, height=240,units='mm', dpi=300)

```

```{r}
# Use the ones not merged yet
head(BLM_erbr_lag_scaled)
head(BLM_erbr_scaled)
head(BLM_erbr_newplts)
table(BLM_erbr_lag_scaled$lagsrtstg)

tags2impute <- unique(BLM_erbr_lag_scaled$Tag[BLM_erbr_lag_scaled$lagsrtstg > 0])
length(tags2impute) # 53

# BLM_erbr_newplts[BLM_erbr_newplts$Tag == tags2impute[1],]  ### lost Tag

for(tt in tags2impute){
  
  imputeTag <- BLM_erbr_lag_scaled[BLM_erbr_lag_scaled$Tag == tt,]
  # last stage recorded
  # Last good year data through Years to impute
    years2impute <- imputeTag$Year[imputeTag$lagsrtstg > 0]-1
  # yy <- years2impute[2]  
  for(yy in years2impute) {
    # where stage.x is there, stage.y is missing
    Stage <- rbinom(n = 1, size = 1, 
                    prob = predict(modelimpute,
                                   newdata = BLM_erbr_scaled[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy,],
                                   type="response"))
    BLM_erbr_lag_scaled$Stage[BLM_erbr_lag_scaled$Tag == tt & BLM_erbr_lag_scaled$Year == yy+1] <- Stage
    # BLM_erbr_newplts$Stage[BLM_erbr_newplts$Tag == tt & BLM_erbr_newplts$Year == yy+1] <- Stage
    BLM_erbr_scaled$Stage.y[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy] <-
      BLM_erbr_scaled$Stage.x[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy+1] <- Stage
    
    if(Stage == 1){
      BLM_erbr_lag_scaled$StageName[BLM_erbr_lag_scaled$Tag == tt & BLM_erbr_lag_scaled$Year == yy+1] <-
        # BLM_erbr_newplts$StageName[BLM_erbr_newplts$Tag == tt & BLM_erbr_newplts$Year == yy+1] <-
        BLM_erbr_scaled$StageName.y[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy] <-
        BLM_erbr_scaled$StageName.x[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy+1] <- "reproductive"
    } else {
      # BLM_erbr_newplts$StageName[BLM_erbr_newplts$Tag == tt & BLM_erbr_newplts$Year == yy+1] <-
        BLM_erbr_lag_scaled$StageName[BLM_erbr_lag_scaled$Tag == tt & BLM_erbr_lag_scaled$Year == yy+1] <- 
        BLM_erbr_scaled$StageName.y[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy] <-
        BLM_erbr_scaled$StageName.x[BLM_erbr_scaled$Tag == tt & BLM_erbr_scaled$Year.x == yy+1] <- "vegetative"
    }
  } # end sequential imputing
}

# table(BLM_erbr2$Stage.x, BLM_erbr2$Stage.y) 
table(BLM_erbr_scaled$Stage.x, BLM_erbr_scaled$Stage.y) 
table(BLM_erbr_scaled$StageName.y, BLM_erbr_scaled$StageName.x) # one dead to seedling?? No, NA BB first year
BLM_erbr_scaled[BLM_erbr_scaled$StageName.y == "seedling",] ## for "BB_7_153"
BLM_erbr_lag[BLM_erbr_lag$Tag == "BB_7_153",]

BLM_erbr_scaled$StageName.x <- factor(BLM_erbr_scaled$StageName.x, 
                                      levels = c("seedling","vegetative","reproductive","dead"), ordered = TRUE)
BLM_erbr_scaled$StageName.y <- factor(BLM_erbr_scaled$StageName.y, 
                                      levels = c("seedling","vegetative","reproductive","dead"), ordered = TRUE)

BLM_erbr_scaled[BLM_erbr_scaled$StageName.y == "seedling",]
# Fate, Stage
table(BLM_erbr_scaled$StageName.y, BLM_erbr_scaled$StageName.x)

# table(BLM_erbr_newplts$StageName)

```





Use popbio to construct stage-structured projection matrices from transition tables
```{r}


BLM_erbr2 <- BLM_erbr_scaled
head(BLM_erbr2) # now imputed missing data
head(BLM_erbr_newplts)
BLM_erbr2 <- merge(BLM_erbr2, BLM_erbr_newplts, # [!duplicated(BLM_erbr_newplts[,c(1:4,9,23:24)]),c(1:4,9,23:24)], 
                   by.x = c("SiteName","Site","Year.x"), 
                   by.y = c("SiteName","Site","Year"))

BLM_MPM <- BLM_erbr2 
names(BLM_MPM) <- c("SiteName","Site","Year","Tag","StageNum","Stage",
                    "surv","lag","lagsurv", 
                    "meanTmean.summer.x","meanTmean.fall.x","meanTmean.winter.x",
                    "TotPpt.summer.x","TotPpt.fall.x","TotPpt.winter.x",
                    # names(BLM_MPM)[9:14],
                    "Year1","FateNum","Fate","surv1",
                    "lag1","lagsurv1", 
                    "meanTmean.summer.y", "meanTmean.fall.y", "meanTmean.winter.y", 
                    "TotPpt.summer.y","TotPpt.fall.y","TotPpt.winter.y",
                    "Year.num","TransectNew.num","num.newPlts", "num.newPlts1")
                    # names(BLM_MPM)[20:29])

# Add anonymous reproduction based on the proportional reproductive outputs of flowering plants and total number of seedlings at the end of the projection interval 
# BLM_newplts <- read.csv("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/ErBr_scripts_May2021/erbr_BLM_newplts.csv")
# BLM_MPM <- merge(BLM_MPM, BLM_newplts, by = c("Year","SiteName","Site"))

# Assign random number of fruit per reproductive individual; NAs will produce errors
BLM_MPM$fruit <- sapply(1:nrow(BLM_MPM), function(rw){
  if(BLM_MPM$StageNum[rw] == 1 && !is.na(BLM_MPM$StageNum[rw])){
    # From DBG, amount of fruit depends on Summer precipitation
    max(1, rpois(1,20+(10*BLM_MPM$TotPpt.summer.x[rw]))) 
    } else {
      0
      }
  })

hist(BLM_MPM$fruit[BLM_MPM$StageNum == 1])
table(BLM_MPM$fruit)

ggplot(BLM_MPM[BLM_MPM$StageNum == 1,], aes(fruit))+
  geom_histogram()+
  theme_bw()

BLM_MPM[is.na(BLM_MPM$fruit) == TRUE,] # BLM_MPM$StageNum == 1 & 
max(BLM_MPM$fruit)

head(BLM_MPM)
# if vegetative then no fruit
table(BLM_MPM$StageNum, BLM_MPM$fruit)

ggplot(BLM_MPM[BLM_MPM$fruit > 0,], aes(TotPpt.summer.x, fruit, colour = SiteName))+
  geom_jitter(height = 0, width = 0.2, alpha = 0.2)+
  stat_smooth(method = "lm")+
  xlab("Summer Precipitation")


## count reproductive plants in year t and seedlings in year t+1
# splitdf <- split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE) 
# x <- splitdf[[35]] # 2 new plants
# x <- splitdf[[119]]

# BLM_erbr2$Tag[BLM_erbr2$StageName.x == "dead" & BLM_erbr2$StageName.y != "dead"]
# BLM_erbr2[BLM_erbr2$Tag == "BB_10_152",]

# BLM_reproplts <- lapply(split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE ), function(x){
BLM_reproplts <- lapply(split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE ), function(x){
  repro <- sum(x$StageNum, na.rm = TRUE)
  if(is.na(x$num.newPlts1[1]) ==  TRUE){
    x$seedling <- 0
  } else {
    x$seedling <- x$fruit / sum(x$fruit, na.rm = TRUE)*x$num.newPlts1[1] # anonymous contribution
  }
  
  # transition frequency table for T matrix
  tf <- table(x$Fate, x$Stage)
  T.mat <- prop.table(tf,2)[-4,-4]  # [-3,-3] # remove "dead" fate and stage
  T.mat[is.nan(T.mat)] <- 0
  
  fert <- tapply(x$seedling, x$Stage, mean)[-4]
  # print(fert)
  fert[is.na(fert)] <- 0
  F.mat <- T.mat * 0
  F.mat[1,] <- fert
  
  A.mat <- T.mat + F.mat
  A.mat

})

st <- "CG"
CI_BLM <- do.call(rbind,lapply(c("CG","DGBLM","GPQ","BB"), function(st){ 
  stout <- stoch.growth.rate(BLM_reproplts[grep(st, names(BLM_reproplts))])
  out <- data.frame(Site = st, approx_lmda = exp(stout$approx),
                    sim_lmda = exp(stout$sim),
                    lwCI = exp(stout$sim.CI[1]),
                    upCI = exp(stout$sim.CI[2]))
  out
  }))

lapply(BLM_reproplts[grep("BLM", names(BLM_reproplts))], function(x) lambda(x))

lapply(BLM_reproplts[grep("BLM", names(BLM_reproplts))], function(x) net.reproductive.rate(x))


x <- split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE)[[2]]

BLM_Amat_CI <- do.call(rbind,lapply(split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE ), function(x){
  ## resampling
  res_l <- unlist(lapply(1:100, function(i){
    x_s <- sample(nrow(x), replace = TRUE)
    bt <- x[x_s,]
    if(is.na(bt$num.newPlts1[1]) ==  TRUE){
      bt$seedling <- 0
      } else {
        bt$seedling <- bt$fruit / sum(bt$fruit, na.rm = TRUE)*bt$num.newPlts1[1] # anonymous contribution
        }
    pro_m <- projection.matrix(bt, stage = "Stage", fate = "Fate", fertility = "fruit")[-3,-3]
    lambda(pro_m)
    }))
  ci <- quantile(res_l, c(0.025, 0.975)) # 95%CI
  data.frame(x[1,c(1:3,9:15,20:29)], lwL = ci[1], upL = ci[2], lambda = res_l)

}))

BLM_Amat_CI$SiteName <- factor(BLM_Amat_CI$SiteName, 
                               levels = c("CastleGarden_BLM", "Droney_BLM", "GardenParkQuarry_BLM", "BigBend_BLM"))

ggsave("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2020_Eriogonum-brandegeei_AprilGoebl_PVA/2021_Eriogonum-brandegeei_integrated_figures/BootstrapResampling.jpg",
    ggplot(BLM_Amat_CI, aes(Site, lambda, group = Site))+
      facet_wrap(~SiteName + Year, scales = "free_x", ncol = 4)+
      geom_jitter(width = 0.05, alpha = 0.5)+
      theme_bw()+
      # geom_violin()+
      geom_boxplot()+
      geom_errorbar(aes(ymin = lwL, ymax = upL))+
      scale_x_discrete(labels =  NULL, breaks = NULL)+  # (guide = guide_axis(angle=90))+
      geom_hline(yintercept = 1, col = "red")+
      xlab(""),
width = 240, height = 120, units = 'mm', dpi=300)


```

Bootstrap observed census transitions
```{r}
x <- split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE)[[20]]



BLM_stocklmbd <- do.call(rbind,lapply(split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE ), function(x){
  # if(is.na(x$num.newPlts1[1]) ==  TRUE){
  #   x$seedling <- 0
  # } else {
  #   x$seedling <- x$fruit / sum(x$fruit, na.rm = TRUE)*x$num.newPlts1[1] # anonymous contribution
  # }

  # x_boot <- x[,c("Stage","Fate","fruit","seedling")]
  # names(x_boot) <- c("stage","fate","fruits","seedling")
  # boot_t <- boot.transitions(x_boot, 100)
  # ci <- quantile(boot_t$lambda, c(0.025, 0.975)) # 95%CI
  # data.frame(x[1,c(1:3,9:15,20:29)], lwL = ci[1], upL = ci[2], lambda = boot_t$lambda)
  
  ## without fertility since those aren't for sure
  boot_t <- boot.transitions(x, 100, stage = "Stage", fate = "Fate")
  ci <- quantile(boot_t$lambda, c(0.025, 0.975))
  data.frame(x[1,c(1:3,9:15,20:29)], lwL = ci[1], upL = ci[2], lambda = boot_t$lambda)
}))


table(BLM_stocklmbd$Site)
ggplot(BLM_stocklmbd, aes(Site, lambda, group = interaction(SiteName, Year)))+
  geom_jitter(width = 0.02, alpha = 0.5)+
  geom_boxplot()+
  theme_bw()+
  facet_wrap(~SiteName + Year, scales = "free_x")

```

Devenish Nelson et al 2010  
```{r}
BLM_stocklmbd <- do.call(rbind,lapply(split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE ), function(x){
  if(is.na(x$num.newPlts1[1]) ==  TRUE){
    x$seedling <- 0
  } else {
    x$seedling <- x$fruit / sum(x$fruit, na.rm = TRUE)*x$num.newPlts1[1] # anonymous contribution
  }
  
  surv <- dbinom(x =  length(x$Fate[x$Fate != "dead"]),  # events or successes, survived what rate to get that 
                 size = nrow(x), prob = length(x$Fate[x$Fate != "dead"])/nrow(x))
  veg <- x[x$Stage == "vegetative",]
  surv_veg <- dbinom(x = length(veg$Fate[veg$Fate == "vegetative"]),
                     size = nrow(veg),
                     prob = length(veg$Fate[veg$Fate == "vegetative"])/nrow(veg))
  rep <- x[x$Stage == "reproductive",]
  surv_rep <- dbinom(x = length(rep$Fate[rep$Fate == "repetative"]),
                     size = nrow(rep),
                     prob = length(rep$Fate[rep$Fate == "repetative"])/nrow(rep))
  # 
  
  
  #Shifted point estimate for average seedlings removing possibility of zero seed for reproductive individuals 
  fecund <- 
  
  data.frame(x[1,c(1:3,9:15,20:29)], lwL = ci[1], upL = ci[2], lambda = boot_t$lambda)
  
}))


```



```{r}
# x <- split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE )[[1]]
BLM_count <- do.call(rbind,lapply(split(BLM_MPM, list(BLM_MPM$Site, BLM_MPM$Year), drop = TRUE ), function(x){
  repro <- sum(x$StageNum, na.rm = TRUE)
  veg <- nrow(x[x$StageNum != 1,])
  total <- nrow(x)
  data.frame(reproNum = repro, vegNum = veg, totalCount = total, Year = x$Year[1], SiteName = x$SiteName[1], Transect = x$Site[1])
}))

ggplot(WD, aes(Year, Water.Deficit, colour = Center))+
  geom_point()+
  geom_line()+
  facet_wrap(~Months, ncol = 1)+
  theme_bw()

table(BLM_MPM$Site)
head(BLM_reproplts)
strsplit(names(BLM_reproplts)[1], "[.]")


# What happened to each transect?
lambdaCD <- data.frame(lmbd = unlist(lapply(BLM_reproplts, function(mat) lambda(mat))),
                       Transect = unlist(lapply(names(BLM_reproplts), function(st) strsplit(st, "[.]")[[1]][1])),
                       Year = as.numeric(unlist(lapply(names(BLM_reproplts), function(st) strsplit(st, "[.]")[[1]][2]))),
                       Year1 = as.numeric(unlist(lapply(names(BLM_reproplts), function(st) strsplit(st, "[.]")[[1]][2])))+1) 

# if each lambda is the the point estimate, events are the number of individuals in time t+1 - the ones that survived, and trials are time t
BLM_CI <- merge(lambdaCD, BLM_count, by = c("Year","Transect"))
BLM_CI <- merge(BLM_CI, BLM_count, by.x = c("Year1", "Transect"), by.y = c("Year", "Transect"))
head(BLM_CI)

lapply(1:nrow(BLM_CI), function(x) dbinom(x = BLM_CI$totalCount.y[x], size = BLM_CI$totalCount.x[x], prob = BLM_CI$lmbd))

# Want climate resulting in the second year of the transition
lambdaCD$Center1 <- "SalidaNorth" # Big Bend and Droney
lambdaCD$Center1[grep("GPQ", lambdaCD$Transect)] <- "CanonCity"
lambdaCD$Center1[grep("CG", lambdaCD$Transect)] <- "SalidaSouth" 
table(lambdaCD$Center1)
# Castle Garden is in the southern Salida subunit and Marsh Quarry and Oil Well are in the Canon City unit.  
# lambdaCD$Center[grep("CG|GPQ" ,lambdaCD$Transect)] <- "CanonCity"
# IF lump by site, not transect
# lambdaCD$Center[grep("CastleGarden", lambdaCD$Transect)] <- "SalidaSouth"
# lambdaCD$Center[grep("CG", lambdaCD$Transect)] <- "SalidaSouth"

# lambdaCD$Center[grep("CastleGarden|GardenParkQuarry" ,lambdaCD$Transect)] <- "CanonCity"
# lambdaCD$Center[grep("GardenParkQuarry" ,lambdaCD$Transect)] <- "CanonCity"
lambdaCD$Center <- "Salida"
lambdaCD$Center[grep("GPQ" ,lambdaCD$Transect)] <- "CanonCity"
table(lambdaCD$Center, lambdaCD$Transect)
table(lambdaCD$Center)
table(lambdaCD$Center1)

lambdaCD <- merge(lambdaCD, BLM_count, by = c("Year","Transect"))

# lamsort <- with(lambdaCD, reorder(Transect, lmbd, median, order = TRUE))
# lambdaCD$Transect <- factor(lambdaCD$Transect, levels = levels(lamsort))

# or sort by names
# lambdaCD$Transect <- factor(lambdaCD$Transect, levels = levels(transLevels))

datetoday <- as.character(Sys.Date())
ggsave(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2021_Eriogonum-brandegeei_SSA/",datetoday,"_BLM_Eriogonum-brandegeei_lambdas.jpg", sep=""),
    # ggplot(lambdaCD, aes(lmbd, Transect))+
    ggplot(lambdaCD, aes(Center1,lmbd, colour = SiteName, group = SiteName))+
      geom_hline(yintercept = 1, color = "red", width = 2)+
      geom_boxplot()+
      geom_jitter(width = 0.2)+  #(height = 0.4) + #, aes(colour = as.factor(Year)), shape = 17)+
      # scale_colour_discrete(name = "Year")+
      # facet_wrap(~Center1+Year, scales = "free")+ 
      facet_wrap(~Year + Center1, scales = "free_x", ncol = 3)+ 
      # theme(axis.text.x=element_text(angle = -90, hjust = 0))+
      theme_bw()+
      ylab(expression(lambda))+
  scale_x_discrete(labels = NULL, breaks = NULL),
width=250, height=270,units='mm', dpi=300)


## CI that the random interval will contain the mean 95% of the time
sites <- c("GPQ","BB","CG","DG")
CIs <- do.call(rbind, lapply(sites, function(x){
    z <- qnorm((1-.95)/2, lower.tail = FALSE)
    xbar <- mean(lambdaCD$lmbd[grep(x, lambdaCD$Transect)])
    sdx <- sqrt(var(lambdaCD$lmbd[grep(x, lambdaCD$Transect)])/nrow(lambdaCD[grep(x, lambdaCD$Transect),]))
    data.frame(Site = x, meanLmbd = xbar, lwCI = xbar - z*sdx, upCI = xbar + z*sdx)
  }))
  
AUs <- names(table(lambdaCD$Center1))
CIs_AU <- do.call(rbind, lapply(AUs, function(x){
    z <- qnorm((1-.95)/2, lower.tail = FALSE)
    xbar <- mean(lambdaCD$lmbd[grep(x, lambdaCD$Center1)])
    sdx <- sqrt(var(lambdaCD$lmbd[grep(x, lambdaCD$Center1)])/nrow(lambdaCD[grep(x, lambdaCD$Center1),]))
    data.frame(Site = x, meanLmbd = xbar, lwCI = xbar - z*sdx, upCI = xbar + z*sdx)
  }))
  

ggsave(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2021_Eriogonum-brandegeei_SSA/",datetoday,"_BLMbyYear_Eriogonum-brandegeei_lambdas.jpg", sep=""),
ggplot(lambdaCD, aes(Transect,lmbd, colour = Transect))+
  geom_hline(yintercept = 1, color = "red", width = 2)+
  # geom_boxplot()+
  # geom_jitter(width = 0.2)+  #(height = 0.4) + #, aes(colour = as.factor(Year)), shape = 17)+
  geom_point()+
  # scale_colour_discrete(name = "Year")+
  # facet_wrap(~Center1+Year, scales = "free")+ 
  facet_wrap(~Year + Center1, scales = "free_x", ncol = 3)+ 
  # theme(axis.text.x=element_text(angle = -90, hjust = 0))+
  theme_bw()+
  ylab(expression(lambda))+
  xlab("Transect")+
  scale_x_discrete(labels = NULL, breaks = NULL)+
  scale_colour_discrete(guide = FALSE),
width=150, height=170,units='mm', dpi=300)



table(names(BLM_reproplts))
table(lambdaCD$Transect, lambdaCD$Center)

table( lambdaCD$Year,lambdaCD$Transect)

# Canon city is GPQ 
length(grep("GPQ", names(BLM_reproplts)))/4
# Salida North is DG and BB
length(grep("DG", names(BLM_reproplts)))/4
length(grep("BB", names(BLM_reproplts)))/2
# Salida South is CG
length(grep("CG", names(BLM_reproplts)))/4
# 
# lambdaCD <- merge(lambdaCD, WD, by.x = c("Year","Center1"), by.y = c("Year","Center"))
# lambdaCD <- merge(lambdaCD, WD, by.x = c("Center1","Months","Year1"), by.y = c("Center","Months","Year"))
lambdaCD <- merge(lambdaCD, WD[WD$Months == "Annual",], by = c("Year", "Center")) # by.x = c("Year","Center1"), by.y = c("Year","Center"))
lambdaCD <- merge(lambdaCD, WD[WD$Months == "Annual",], by.x = c("Center","Months","Year1"), by.y = c("Center","Months","Year"))
head(lambdaCD) # .y is year1, .x is year
table(lambdaCD$Center1);table(lambdaCD$Center)

WD_t <- ggplot(lambdaCD, aes(Water.Deficit.x, lmbd, colour = Center1))+
  geom_jitter(height = 0, width = 0.1)+
  stat_smooth(method = "lm", alpha = 0.3)+
  facet_wrap(~Months, scales = "free")+
  ylab(expression(lambda[t-t+1]))+
  xlab(expression(paste("Water deficit in ", year[t])))+
  theme_bw()
WD_t1 <- ggplot(lambdaCD[lambdaCD$Months != "ASON",], aes(Water.Deficit.y, lmbd, colour = Center1))+
  geom_jitter(height = 0, width = 0.1)+
  stat_smooth(method = "lm", alpha = 0.3)+
  facet_wrap(~Months, scales = "free")+
  ylab(expression(lambda[t-t+1]))+
  xlab(expression(paste("Water deficit in ", year[t+1])))+
  theme_bw()

ggsave(paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Eriogonum brandegeei/2021_Eriogonum-brandegeei_SSA/",datetoday,"_BLM_Eriogonum-brandegeei_lambdasXwaterdeficit.jpg", sep=""),
WD_t/WD_t1,
width=150, height=140,units='mm', dpi=300)
  



summary(lm(log(lmbd) ~ Water.Deficit.x + Center1, data = lambdaCD[lambdaCD$Months == "Annual",]))

plot(lambdaCD$totalCount,scale(lambdaCD$lmbd))
plot(lambdaCD$totalCount, log(lambdaCD$lmbd))
plot(lambdaCD$totalCount, lambdaCD$lmbd)

lambda(mean(BLM_reproplts[grep("BB", names(BLM_reproplts))])) # , na.rm = TRUE
lambda(mean(BLM_reproplts[grep("CG", names(BLM_reproplts))]))
lambda(mean(BLM_reproplts[grep("DGBLM", names(BLM_reproplts))]))
lambda(mean(BLM_reproplts[grep("GPQ", names(BLM_reproplts))]))

trans <- unique(BLM_erbr2$Site)
meanLmbd_trans <- do.call(rbind,lapply(trans, function(x){
  data.frame(Trans = x, lmbd = lambda(mean(BLM_reproplts[grep(x, names(BLM_reproplts))])))
}))

sites <- c("GPQ","BB","CG","DG")
lapply(sites, function(x) sd(meanLmbd_trans$lmbd[grep(x, meanLmbd_trans$Trans)]))

# Want SD across transects averaged over years per transect
sd(unlist(lapply(BLM_reproplts[grep("BB", names(BLM_reproplts))], function(mat){ 
  lambda(mat)
  })))
sd(unlist(lapply(BLM_reproplts[grep("CG", names(BLM_reproplts))], function(mat){ 
  lambda(mat)
  })))
sd(unlist(lapply(BLM_reproplts[grep("DG", names(BLM_reproplts))], function(mat){ 
  lambda(mat)
  })))
sd(unlist(lapply(BLM_reproplts[grep("GPQ", names(BLM_reproplts))], function(mat){ 
  lambda(mat)
  })))

lambda(mean(BLM_reproplts[grep("BigBend", names(BLM_reproplts))])) 
lambda(mean(BLM_reproplts[grep("Droney", names(BLM_reproplts))])) 
lambda(mean(BLM_reproplts[grep("Castle", names(BLM_reproplts))])) 
lambda(mean(BLM_reproplts[grep("GardenPark", names(BLM_reproplts))]))

var(unlist(lapply(BLM_reproplts[grep("BigBend", names(BLM_reproplts))], function(mat) lambda(mat)))) # , na.rm = TRUE
var(unlist(lapply(BLM_reproplts[grep("Droney", names(BLM_reproplts))], function(mat) lambda(mat)))) # , na.rm = TRUE
var(unlist(lapply(BLM_reproplts[grep("CastleGarden", names(BLM_reproplts))], function(mat) lambda(mat)))) # , na.rm = TRUE
var(unlist(lapply(BLM_reproplts[grep("GardenPark", names(BLM_reproplts))], function(mat) lambda(mat)))) # , na.rm = TRUE

# Salida North
lambda(mean(BLM_reproplts[grep("BB|DGBLM", names(BLM_reproplts))])) # , na.rm = TRUE
lambda(mean(BLM_reproplts[grep("CG|GPQ", names(BLM_reproplts))]))

var(unlist(lapply(BLM_reproplts[grep("BB|DGBLM", names(BLM_reproplts))], function(mat) lambda(mat)))) # , na.rm = TRUE
var(unlist(lapply(BLM_reproplts[grep("CG|GPQ", names(BLM_reproplts))], function(mat) lambda(mat))))

## Within year mean and variation across sites in a center
aggregate(lmbd ~ Center1 + Year + Water.Deficit.x, data = lambdaCD, mean)
SD_BLM <- aggregate(lmbd ~ Center1 + Year + Water.Deficit.x, data = lambdaCD, sd)

summary(glm(lmbd ~ Water.Deficit.x, family = quasipoisson(link = "log"), data = SD_BLM))
ggplot(SD_BLM, aes(Water.Deficit.x, lmbd))+
  geom_point()+
  ylab("SD")+
  theme_bw()

```


BLM and DBG together!
```{r}
head(DBG_GP)
str(DBG_GP)
DBG_GP$YEAR <- as.numeric(as.character(DBG_GP$YEAR))
DBG_GP$Center1 <- "CanonCity"
DBG_GP <- merge(DBG_GP, WD[WD$Months == "Annual",], by.x = c("YEAR","Center1"), by.y = c("Year","Center")) 
DBG_GP$Year1 <- DBG_GP$YEAR+1
DBG_GP <- merge(DBG_GP, WD[WD$Months == "Annual",], by.x = c("Center1","Months","Year1"), by.y = c("Center","Months","Year"))

medianLmbd <- aggregate(LAMBDA ~ TRANSECT + YEAR + Year1 + Water.Deficit.x + Water.Deficit.y + Center1, 
                        median, data = DBG_GP)
# medianLmbd <- merge(medianLmbd, WD[WD$Center == "CanonCity" & WD$Months == "Annual",],
#                     by.x = "YEAR", by.y = "Year")


library(boot)
Bmean <- function(DBG_GP, indices) {
  d <- DBG_GP[indices] # allows boot to select sample 
    return(mean(d))
} 
results <- boot(DBG_GP$LAMBDA, Bmean, R = 99, stype = "i")
plot(results)
boot.ci(results, type=c("norm"))

names(lambdaCD);names(medianLmbd)
DBG2bind <- medianLmbd[,c("Center1","YEAR","Year1","TRANSECT","LAMBDA","Water.Deficit.x","Water.Deficit.y")]
BLM2bind <- lambdaCD[,c("Center1","Year","Year1","Transect","lmbd","Water.Deficit.x","Water.Deficit.y")]

names(DBG2bind) <- c("Center1","Year","Year1","Transect","lmbd","Water.Deficit.x","Water.Deficit.y")

lmbd_all <- rbind(DBG2bind,BLM2bind)
table(lmbd_all$Center1)
```


```{r}
aggregate(lmbd ~ Center1 + Year + Water.Deficit.x + Water.Deficit.y, mean, data = lmbd_all)
all_sd <- aggregate(lmbd ~ Center1 + Year + Water.Deficit.x + Water.Deficit.y, sd, data = lmbd_all)

ggplot(lmbd_all, aes(Water.Deficit.x, lmbd, colour = Center1))+
  geom_point()+
  theme_bw()+
  stat_smooth(method = "lm")

ggplot(all_sd, aes(Water.Deficit.x, lmbd, colour = Center1))+
  geom_point()+
  theme_bw()+
  stat_smooth(method = "lm")

## Average lambda by center
## average SD by center (SD within a year over transects)
aggregate(lmbd ~ Center1, lmbd_all, mean)
aggregate(lmbd ~ Center1,all_sd, mean)


```




