---
title: "erbr_IPM_stochastic"
author: "April Goebl"
date: "2/5/2021"
output: html_document
---

```{r setup, include=FALSE}

rm(list=ls())
source('eigenall.r')
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(matrixStats)
library(viridis) 
library("stringr")

```


```{r, echo=FALSE }

## LOAD MATRICES FOR EACH YEAR (STUDY AND NON-STUDY PERIOD) & TRANSECT FROM IPMdeterm CODE 
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_P1k_20230503")  #Output generated from erbr_4IPMdeterm script 

erbr <- read.csv("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/ErBr/erbr_TagClust2022_20230408.csv", header=TRUE)
erbr$Year <- as.factor(erbr$Year)

```



```{r}

## OBTAIN STABLE STAGE STRUCTURE
## Modified from Dan's code 
N.startNum <- 61                  #Set starting pop sz as 2022 obs sz; could increase to better reflect pop sz, not just demog plts in transects

mx.num <- length(mx.list)
mx.mean <- Reduce('+', mx.list)/mx.num  #Calculate mean of all matrices 
mnees=eigenall(mx.mean)
N.vecStart=N.startNum * mnees$stable.stage

```



## STOCHASTIC LAMBDA (SHORT TERM)
```{r}

## Calculate stochastic lambda for each transect separately

names.mx <- unique(names(mx.list))
names.trans <- unique(sub('.*\\_', '', names.mx))     #Keep text following '_'
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
names.param <- unique(sub('\\-.*', '', names.mx))     #Keep text preceding '-'

sim.yrs <- 10  #Number of years to simulate, e.g. 100
reps <- 10     #Number of replicate runs to make for env variability
#For 10x10 took <30 mins to run on my laptop ***

## Variables for output 
stoch.lams.template <- as.data.frame(matrix(NA, nrow=(reps*length(names.param)*length(names.trans)), ncol=4))  
colnames(stoch.lams.template) <- c("PARAM","TRANS", "REP", "LAMBDA")
stoch.lams.template$REP <- rep(1:reps)
stoch.lams.template$TRANS <- rep(names.trans, each=reps)
stoch.lams.template$PARAM <- rep(names.param, each=(nrow(stoch.lams.template)/length(names.param)))

stoch.lams.all <- NULL



## Start parameter loop (for a given parameter draw, select X yrs (e.g. 100))
for (pp in 1:length(names.param)) {
  mx.param <- mx.list[grep(names.param[pp], names(mx.list))]         #Subset mx.list by each parameter
  stoch.lams <- subset(stoch.lams.template, PARAM==names.param[pp])  #Subset output template by param in parameter loop 

  
  ## Start transect loop
  for (tt in 1:length(names.trans)) {
  mx.trans <- mx.param[grep(names.trans[tt], names(mx.param))]       #Subset mx list by each transect
  

  ## Modified from Dan 
  ## Input parameters
  mx.num.trans <- length(mx.trans)               #This includes all yr matrices for a given param & transect
  probs <- rep(1/mx.num.trans, mx.num.trans)     #List of probs of occurrence for the lams: should sum to 1

  ## Make a vector of random values to use to pick between all yr matrices:
  mx.randVals <- matrix(sample(1:mx.num.trans, size=reps*sim.yrs, prob=probs, replace=TRUE), nrow=sim.yrs, ncol=reps)
  ## End of inputs
  

  lams <- matrix(0, sim.yrs-1, reps)                              #Set up a matrix for the pop sizes to fill
  stoch.lams.trans <- subset(stoch.lams, TRANS==names.trans[tt])  #Subset output template by trans in transect loop
  N.vec <- N.vecStart                                             #Re-set to desired starting values

  
  ## Loop over reps and years 
  for (jj in 1:reps) {
    N.vec <- N.vecStart
    
    for (ii in 1:(sim.yrs-1)) {
  
      pick <- mx.randVals[ii,jj]                         #Pick a mx corresponding to 1 yr (i.e. set of clim conditions) 
      mx <- mx.trans[pick]

      oldN.vec <- N.vec
      ## Multiply matrix by vec w num indivs per sz class (initially this is stbl stage struct)
      N.vec <- as.matrix(as_tibble(mx)) %*% N.vec        #Grow the population one year
    
      lams[ii,jj] <- sum(N.vec)/sum(oldN.vec)
      N.vec <- N.startNum*N.vec/sum(N.vec)               #Restandardize the tot nums in vector so doesn't get too big or sm 
    } #End of ii (years) loop
    
    stoch.lams.trans[jj,4]= mean(lams[,jj])                              
  } #End of jj (reps) loop
## End of Dan's code 
  
  stoch.lams.all <- rbind(stoch.lams.all, stoch.lams.trans)
} #End of tt (transect loop)

} #End of pp (parameter) loop

```


```{r}

## SAVE OUTPUT
date <- Sys.Date()        #Get date to be added to file name
date <- str_replace_all(date, "-", "") #library("stringr")?
#date <- as.character(220807) 
saveRDS(stoch.lams.all, file=paste("erbrLamsStoch_noYRE_y100r100P1k", date, sep="_"))

stoch.lams.all <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrLamsStoch_noYRE_y100r100P1k_220807")



## LOOK AT RESULTS
hist(stoch.lams.all$LAMBDA, breaks=150, border=FALSE, xlab="Short term stochastic population growth rate",
     main=NA, cex.lab=1.4, cex.axis=1.2, xlim=c(0.75,1.05), col="grey90")
abline(v=1, col="red", lwd=0.75)


## Plot transects individually 
cols <- viridis(12)
brks <- c(20,30,40,30,40,40,40,30,40,40,30,40)
par(mfrow=c(3,4))                      

for (tt in 1:length(names.trans)) {
  lam.trns <- subset(stoch.lams.all, TRANS==names.trans[tt])

hist(lam.trns$LAMBDA, breaks=brks[tt], border=FALSE, xlab="Short term stochastic lambda",
     main=names.trans[tt], cex.lab=1.4, cex.axis=1.2, xlim=c(0.6,1.2), col=cols[tt])
abline(v=1, col="red", lwd=0.5)
}


## Calculate mean per transect
lam.mean.trans <- stoch.lams.all %>% group_by(TRANS) %>% summarise(MEAN=mean(LAMBDA), 
                  MED=median(LAMBDA), MAX=max(LAMBDA), MIN=min(LAMBDA))
min(lam.mean.trans$MEAN)
max(lam.mean.trans$MEAN)
mean(stoch.lams.all$LAMBDA)  ## Overall mean

```




## STOCHASTIC LAMBDA (SHORT TERM) FOR EACH DATASET 
```{r}

## ESTIMATE SS_LAMBDA FOR EACH DATASET; 'IGNORE' TRANSECT DIFFERENCES
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_P1k_20230503")  #Output from erbr_4IPMdeterm script
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to13_P1k_20231026") 
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to13even_P1k_20231026") 
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to13odd_P1k_20231026") 
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to8_P1k_20231028") 
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_9to13_P1k_20231028") 


## OBTAIN STABLE STAGE STRUCTURE
## Modified from Dan's code 
N.startNum <- 61                  #Set starting pop sz as 2022 obs sz; could increase to better reflect pop sz, not just demog plts in transects

mx.num <- length(mx.list)
mx.mean <- Reduce('+', mx.list)/mx.num  #Calculate mean of all matrices 
mnees=eigenall(mx.mean)
N.vecStart=N.startNum * mnees$stable.stage


names.mx <- unique(names(mx.list))
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
names.param <- unique(sub('\\-.*', '', names.mx))     #Keep text preceding '-'

sim.yrs <- 100  #Number of years to simulate, e.g. 100
reps <- 100     #Number of replicate runs to make for env variability, e.g. 100

## Variables for output 
stoch.lams.template <- as.data.frame(matrix(NA, nrow=(reps*length(names.param)), ncol=3))  
colnames(stoch.lams.template) <- c("PARAM", "REP", "LAMBDA")
stoch.lams.template$REP <- rep(1:reps)
#stoch.lams.template$TRANS <- rep(names.trans, each=reps)
stoch.lams.template$PARAM <- rep(names.param, each=(nrow(stoch.lams.template)/length(names.param)))

stoch.lams.all <- NULL


## Start parameter loop (for a given parameter draw, select X yrs (e.g. 100))
for (pp in 1:length(names.param)) {
  mx.param <- mx.list[grep(names.param[pp], names(mx.list))]         #Subset mx.list by each parameter
  stoch.lams <- subset(stoch.lams.template, PARAM==names.param[pp])  #Subset output template by param in parameter loop 
  
  ## Start transect loop
  #for (tt in 1:length(names.trans)) {
  #mx.trans <- mx.param[grep(names.trans[tt], names(mx.param))]       #Subset mx list by each transect
  

  ## Modified from Dan 
  ## Input parameters
  mx.num.param <- length(mx.param)               #This includes all yr matrices for a given param
  probs <- rep(1/mx.num.param, mx.num.param)     #List of probs of occurrence for the lams: should sum to 1

  ## Make a vector of random values to use to pick between all yr matrices:
  mx.randVals <- matrix(sample(1:mx.num.param, size=reps*sim.yrs, prob=probs, replace=TRUE), nrow=sim.yrs, ncol=reps)
  ## End of inputs
  

  lams <- matrix(0, sim.yrs-1, reps)                              #Set up a matrix for the pop sizes to fill
  #stoch.lams.trans <- subset(stoch.lams, TRANS==names.trans[tt])  #Subset output template by trans in transect loop
  N.vec <- N.vecStart                                             #Re-set to desired starting values

  
  ## Loop over reps and years 
  for (jj in 1:reps) {
    N.vec <- N.vecStart
    
    for (ii in 1:(sim.yrs-1)) {
  
      pick <- mx.randVals[ii,jj]                         #Pick a mx corresponding to 1 yr (i.e. set of clim conditions) 
      mx <- mx.param[pick]

      oldN.vec <- N.vec
      ## Multiply matrix by vec w num indivs per sz class (initially this is stbl stage struct)
      N.vec <- as.matrix(as_tibble(mx)) %*% N.vec        #Grow the population one year
    
      lams[ii,jj] <- sum(N.vec)/sum(oldN.vec)
      N.vec <- N.startNum*N.vec/sum(N.vec)               #Restandardize the tot nums in vector so doesn't get too big or sm 
    } #End of ii (years) loop
    
    stoch.lams[jj,3]= mean(lams[,jj])                              
  } #End of jj (reps) loop
## End of Dan's code 
  
  stoch.lams.all <- rbind(stoch.lams.all, stoch.lams)
#} #End of tt (transect loop)

} #End of pp (parameter) loop
## ---------------------------


## SAVE OUTPUT
date <- Sys.Date()        #Get date to be added to file name
date <- str_replace_all(date, "-", "") #library("stringr")?
saveRDS(stoch.lams.all, file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrLamsStoch_noYRE_y100r100P1k_noTrans_9to13", date, sep="_"))
#stoch.lams.all <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrLamsStoch_noYRE_y100r100P1k_220807")



## LOOK AT RESULTS
hist(stoch.lams.all$LAMBDA, breaks=150, border=FALSE, xlab="Short term stochastic population growth rate",
     main=NA, cex.lab=1.4, cex.axis=1.2, xlim=c(0,15), col="grey90")
abline(v=1, col="red", lwd=0.75)


## Plot each dataset in a different panel
#cols <- viridis(12)
#brks <- c(20,30,40,30,40,40,40,30,40,40,30,40)
#par(mfrow=c(3,4))                      

#for (tt in 1:length(names.trans)) {
#  lam.trns <- subset(stoch.lams.all, TRANS==names.trans[tt])

#hist(lam.trns$LAMBDA, breaks=brks[tt], border=FALSE, xlab="Short term stochastic lambda",
#     main=names.trans[tt], cex.lab=1.4, cex.axis=1.2, xlim=c(0.6,1.2), col=cols[tt])
#abline(v=1, col="red", lwd=0.5)
#}


## Calculate mean
mean(stoch.lams.all$LAMBDA)
#lam.mean.trans <- stoch.lams.all %>% group_by(TRANS) %>% summarise(MEAN=mean(LAMBDA), 
#                  MED=median(LAMBDA), MAX=max(LAMBDA), MIN=min(LAMBDA))
#min(lam.mean.trans$MEAN)
#max(lam.mean.trans$MEAN)
#mean(stoch.lams.all$LAMBDA)  ## Overall mean

```




## PROBABILITY OF QUASI-EXTINCTION 
```{r}

## Calculate risk of extinction for each transect separately

## ** Coding should be improved- currently takes a LONG time to run, very inefficient **


names.mx <- unique(names(mx.list))
names.trans <- unique(sub('.*\\_', '', names.mx))     #Keep text following '_'
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
names.param <- unique(sub('\\-.*', '', names.mx))     #Keep text preceding '-'

sim.yrs <- 100    #Num of years to simulate, e.g. 100
reps <- 10       #Num of reps runs to make. How many out of reps does pop fall below thresh? e.g. 500


## Variables for output 
column.names <- c("PARAM","TRANS", "REP", "YEAR", "POP_SZ")
pop.sz.template <- as.data.frame(matrix(NA, nrow=(length(names.param)*reps*sim.yrs*length(names.trans)),
                   ncol=length(column.names)))
colnames(pop.sz.template) <- column.names
pop.sz.template$YEAR <- rep(1:sim.yrs)
pop.sz.template$REP <- rep(1:reps, each=sim.yrs)
pop.sz.template$TRANS <- rep(names.trans, each=sim.yrs*reps)
pop.sz.template$PARAM <- rep(names.param, each=(nrow(pop.sz.template)/length(names.param)))

pop.sz.all <- NULL        



## Start parameter loop (for a given parameter draw, select X yrs)
for (pp in 1:length(names.param)) {

  mx.param <- mx.list[grep(names.param[pp], names(mx.list))]         #Subset mx.list by each parameter
  pop.sz <- subset(pop.sz.template, PARAM==names.param[pp])          #Subset output template by param in parameter loop 

  ## Start transect loop
  for (tt in 1:length(names.trans)) {
      mx.trans <- mx.param[grep(names.trans[tt], names(mx.param))]   #Subset mx list by each transect

    
      ## Modified from Dan 
      ## Input parameters
      mx.num.trans <- length(mx.trans)               
      probs <- rep(1/mx.num.trans, mx.num.trans)      #List of probs of occurrence for the lams: should sum to one
      
      ## Make a matrix of random values to use to pick between all yr matrices:
      mx.randVals <- matrix(sample(1:mx.num.trans, size=reps*sim.yrs, prob=probs, replace=TRUE), nrow=sim.yrs, ncol=reps)
      ## End of inputs
      
      
      pop.sz.trans <- subset(pop.sz, TRANS==names.trans[tt]) #Subset output template by trans in transect loop
      N.vec <- N.vecStart                                    #Re-set to desired starting values
    
      
      ## Loop over reps and years
    for (jj in 1:reps) {
      N.vec <- N.vecStart
      pop.sz.rep <- subset(pop.sz.trans, REP==as.vector(1:reps)[jj])  #Subset output template by rep in rep loop

          for (ii in 1:(sim.yrs)) {
      
          pick <- mx.randVals[ii,jj]                    #Pick a mx corresponding to 1 yr (i.e. set of clim conditions) 
          mx <- mx.trans[pick]
        
          ## Multiply matrix by vec w num indivs per sz class (initially this is stbl stage struct)
          N.vec <- as.matrix(as_tibble(mx)) %*% N.vec   #Grow the population one year
          pop.sz.rep[ii,5]= sum(N.vec)  
        } #End of ii (years) loop
      
      pop.sz.all <- rbind(pop.sz.all, pop.sz.rep)

    } #End of jj (rep) loop
      
  } #End of tt (transect) loop
      
} #End of pp (param) loop




## SAVE OUTPUT
date <- as.character(220808) 
saveRDS(pop.sz.all, file=paste("erbrPopSz_y100r50P250", date, sep="_"))

```


```{r}

## LOOK AT RESULTS

pop.sz.all <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSz_y100r50P250_220808")


## Plot cumulative extinction curves
thresh <- 10  #5 

pop.sz.forPlot <- pop.sz.all %>% tidyr::spread(REP, POP_SZ)
pop.sz.forPlot$PROB <- NA

for (bb in 1:nrow(pop.sz.forPlot)) {
  bel.thresh <- pop.sz.forPlot[bb,4:(reps+3)] < thresh #Check num reps w pop sz < thresh in each yr
  pop.sz.forPlot$PROB[bb] <- (length(bel.thresh[bel.thresh==TRUE])) / reps
}


## Average of all params 
mean.ext <- pop.sz.forPlot %>% group_by(YEAR) %>% summarise(MEAN=mean(PROB))
plot(mean.ext$YEAR, mean.ext$MEAN, type='l', xlab="Year", ylab="Probability of quasi-extinction", ylim=c(0,1))




## Subset by transect
## Plot results including all param uncertainty
cols <- viridis(12)
par(mfrow=c(4,3))  
par(mar=c(4,6,3,2)) 
for (nn in 1:length(names.trans)) {
  trans <- subset(pop.sz.forPlot, TRANS==names.trans[nn])
  plot(trans$YEAR, trans$PROB, type='l', xlab="Year", ylab="Probability of\nquasi-extinction",
  lwd=1.5, col=cols[nn], cex.lab=1.5, cex.axis=1.2, main=names.trans[nn], ylim=c(0,1))
}


## Average of all params 
par(mfrow=c(4,3))  
par(mar=c(4,6,3,2)) 
for (nn in 1:length(names.trans)) {
  trans <- subset(pop.sz.forPlot, TRANS==names.trans[nn])
  mean.trans <- trans %>% group_by(YEAR) %>% summarise(MEAN=mean(PROB))
  plot(mean.trans$YEAR, mean.trans$MEAN, type='l', xlab="Year", ylab="Probability of\nquasi-extinction",
  lwd=2.5, col=cols[nn], cex.lab=1.5, cex.axis=1.2, main=names.trans[nn], ylim=c(0,1))
}

## ** Add CIs on param avg ** 

```





## ESTIMATE POPULATION SIZE PER YEAR & COMPARE TO OBSERVED 
```{r}

## Load matrices

#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_221129")
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_221129")

mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_4to8_210511")
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_9to13_210511")
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_4to13_210511")
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_4to13even_210511")
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_4to13odd_210511")
#mx.list4to8 <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to8_P1k_20231028")
#mx.list9to13 <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_9to13_P1k_20231028")


names.mx <- unique(names(mx.list))
names.trans <- unique(sub('.*\\_', '', names.mx))     #Keep text following '_'
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
names.param <- unique(sub('\\-.*', '', names.mx))     #Keep text preceding '-'


## -- OPTIONAL ------   Subset to keep only years from study period (2004-2017 or longer) 
match.yr <- as.character(c(2004:2021))
mx.list <- mx.list[grep(paste(match.yr,collapse="|"), names(mx.list))]
names.mx <- unique(names(mx.list))
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
## ---------------------------------------------------------------------------

sim.yrs <- length(names.yr)


## OBTAIN STABLE STAGE STRUCTURE
## AG modified from Dan's code 
N.startNum <- 148                         #Set as observed population size in 2004

mx.num <- length(mx.list)
mx.mean <- Reduce('+', mx.list)/mx.num    #Calculate mean of all matrices 
mnees=eigenall(mx.mean)
N.vecStart=N.startNum * mnees$stable.stage


pop.sz.all <- as.data.frame(matrix(NA, nrow=(length(names.param)), ncol=sim.yrs+2))
#colnames(pop.sz.all) <- c("PARAM", as.character(2004:2018))
colnames(pop.sz.all) <- c("PARAM", as.character(2004:2022))
pop.sz.all$PARAM <- rep(names.param)


## Start parameter loop (for a given parameter draw, loop over data yrs)
for (pp in 1:length(names.param)) {
  mx.param <- mx.list[grep(paste(names.param[pp],"-", sep=""), names(mx.list))]      #Subset mx.list by each parameter
  pop.sz <- matrix(NA, nrow=(sim.yrs), ncol=1)


  ## Modified from Dan
  N.vec <- N.vecStart                                               #Re-set to desired starting values

  ## Loop over years (from 2004-2017 to give estimates for 2005-2018)
    for (ii in 1:(sim.yrs)) {
  
      mx.yr <- mx.param[grep(names.yr[ii], names(mx.param))]        #Subset mx list of each param by yr
      mx <- Reduce('+', mx.yr)/length(names.trans)                  #Calculate mean mx of all transects 


      ## Multiply matrix by vec w num indivs per sz class (initially this is stbl stage struct)
      N.vec <- as.matrix(as_tibble(mx)) %*% N.vec        #Grow the population one year
      pop.sz[ii]= sum(N.vec)             
    } # end of ii (years) loop

 pop.sz.all[pp,3:ncol(pop.sz.all)] <- pop.sz
} #End pp loop




## Add observed values for 2004
pop.sz.all$`2004` <- N.startNum


## Calculate mean and SE of predicted pop size
pop.sz.means <- colMeans(pop.sz.all[2:(length(names.yr)+2)])
pop.sz.sds <- matrixStats::colSds(as.matrix(pop.sz.all[2:(length(names.yr)+2)]))

#pop.sz.ciHi <- 0
#pop.sz.ciLo <- 0
#for (cc in 3:ncol(pop.sz.all)) {
#  cis <- t.test(pop.sz.all[,cc], conf.level=0.95)$conf.int
#  pop.sz.ciHi <- c(pop.sz.ciHi, cis[2])
#  pop.sz.ciLo <- c(pop.sz.ciLo, cis[1])
#}

# Calculate standard error of mean 
#calcSE <- function(x){sd(x)/sqrt(length(x))}





## SAVE FOR PLOTTING BELOW BY RENAMING BASED ON DATA SET 
pop.sz.means.full <- pop.sz.means
pop.sz.sds.full <- pop.sz.sds
names.yr.full <- names.yr
pop.sz.ciHi.full <- pop.sz.ciHi
pop.sz.ciLo.full <- pop.sz.ciLo

pop.sz.means.evry <- pop.sz.means
pop.sz.sds.evry <- pop.sz.sds
names.yr.evry <- names.yr
pop.sz.ciHi.evry <- pop.sz.ciHi
pop.sz.ciLo.evry <- pop.sz.ciLo

pop.sz.means.prun <- pop.sz.means
pop.sz.sds.prun <- pop.sz.sds
names.yr.prun <- names.yr
pop.sz.ciHi.prun <- pop.sz.ciHi
pop.sz.ciLo.prun <- pop.sz.ciLo

pop.sz.means.odd <- pop.sz.means
pop.sz.sds.odd <- pop.sz.sds
names.yr.odd <- names.yr
pop.sz.ciHi.odd <- pop.sz.ciHi
pop.sz.ciLo.odd <- pop.sz.ciLo


## Save output 
pop.sz.comb <- as.data.frame(cbind(pop.sz.means,pop.sz.sds,pop.sz.ciHi,pop.sz.ciLo))
colnames(pop.sz.comb) <- c("Mean","SD","CI_HI", "CI_LO")
write.csv(pop.sz.comb,file="C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/20230725_PopSzFullEsts.csv")

## PLOT PREDICTED VS OBSERVED POP SZ
## Calculate observed: Subset by year and count the number of living plants 
num.indivs <- erbr %>% group_by(Year) %>% summarise(numIndivs=sum(surv, na.rm=TRUE)) 
num.indivs$Year <- as.numeric(as.character(num.indivs$Year))
num.indivs <- num.indivs[which(num.indivs$numIndivs!=0),]


## Plot mean and SD (or CIs)
plot(as.numeric(as.character(num.indivs$Year)), num.indivs$numIndivs, type='l', ylim=c(0,250), 
     ylab="Number of individuals",xlab="Year", cex.lab=1.45, cex.axis=1.3)
points(as.numeric(as.character(num.indivs$Year)), num.indivs$numIndivs, pch=16, cex=0.75)

lines(2004:2022,pop.sz.means.full[1:19],
      col="grey60",lwd=1.9)  #pop.sz.means.full[1:length(pop.sz.means.full)]
arrows(2004:2022,pop.sz.means.full[1:19], 
       2004:2022,pop.sz.means.full[1:19]+pop.sz.sds.full[1:19], 
       length=0.05, angle=90, col="grey60", lwd=0.6)
#arrows(2004:2022,pop.sz.means.full[1:19], 
#       2004:2022,pop.sz.means.full[1:19]+pop.sz.ciHi.full[1:19], 
#       length=0.05, angle=90, col="grey60", lwd=0.6)
arrows(2004:2022,pop.sz.means.full[1:19], 
       2004:2022,pop.sz.means.full[1:19]-pop.sz.sds.full[1:19], 
       length=0.05, angle=90, col="grey60", lwd=0.6)

legend("topright", c("Observed", "Full dataset"), col=c("black","grey60"),lwd=2.95,cex=1,
       horiz=FALSE, bty="y",seg.len=1, xpd=TRUE)


## Add other data sets -------------------------------------
lines(2004:2018, pop.sz.means.odd[1:length(pop.sz.means.odd)],
      col="pink",lwd=1.1)
arrows((2004:2018)+0.15,pop.sz.means.odd[1:length(pop.sz.means.odd)], 
       (2004:2018)+0.15,pop.sz.means.odd[1:length(pop.sz.means.odd)]+pop.sz.ciHi.odd[1:length(pop.sz.means.odd)], 
       length=0.05, angle=90, col="pink", lwd=0.6)
arrows((2004:2018)+0.15,pop.sz.means.odd[1:length(pop.sz.means.odd)], 
       (2004:2018)+0.15,pop.sz.means.odd[1:length(pop.sz.means.odd)]-pop.sz.ciLo.odd[1:length(pop.sz.means.odd)], 
       length=0.05, angle=90, col="pink", lwd=0.6)

lines(2004:2018, pop.sz.means.evry[1:length(pop.sz.means.evry)],
      col="lightskyblue",lwd=1.3)
arrows((2004:2018)-0.1,pop.sz.means.evry[1:length(pop.sz.means.evry)], 
       (2004:2018)-0.1,pop.sz.means.evry[1:length(pop.sz.means.evry)]
       +pop.sz.ciHi.evry[1:length(pop.sz.means.evry)],length=0.05, angle=90, col="lightskyblue", lwd=0.6)
arrows((2004:2018)-0.1,pop.sz.means.evry[1:length(pop.sz.means.evry)], 
       (2004:2018)-0.1,pop.sz.means.evry[1:length(pop.sz.means.evry)]
       -pop.sz.ciLo.evry[1:length(pop.sz.means.evry)],length=0.05, angle=90, col="lightskyblue", lwd=0.6)

lines(2004:2018, pop.sz.means.prun[1:length(pop.sz.means.prun)],
      col="springgreen3",lwd=1.05)
arrows((2004:2018)+0.1,pop.sz.means.prun[1:length(pop.sz.means.prun)], 
       (2004:2018)+0.1,pop.sz.means.prun[1:length(pop.sz.means.prun)]+pop.sz.ciHi.prun[1:length(pop.sz.means.prun)], 
       length=0.05, angle=90, col="springgreen3", lwd=0.6)
arrows((2004:2018)+0.1,pop.sz.means.prun[1:length(pop.sz.means.prun)], 
       (2004:2018)+0.1,pop.sz.means.prun[1:length(pop.sz.means.prun)]-pop.sz.ciLo.prun[1:length(pop.sz.means.prun)], 
       length=0.05, angle=90, col="springgreen3", lwd=0.6)

legend("top", c("Observed", "Full dataset", "Consecutive years", "Even years", "Odd years"), col=c("black","grey60","lightskyblue","springgreen3","pink"),lwd=1.95,cex=1,
       horiz=FALSE, bty="y",seg.len=1, xpd=TRUE)

```




## LOOK AT NUMBERS OF NEW OR SM PLANTS IN EACH YEAR
```{r}

## To understand the pop sz increase ~2007, that is not captured by predictions, plot nums of new plts in each yr (by sz group)

## Make df that will hold data containing new plants
years <- unique(erbr$Year)
years <- years[order(years)]
dats.newPlts <- as.data.frame(rep(unique(erbr$TransectNew), each=length(years)))
colnames(dats.newPlts) <- "TransectNew"
dats.newPlts$Year <- rep(years)

## Identify new plants
newPlts <- erbr %>% group_by(TagNew) %>% slice(which.min(Year))  #Identify rows with 1st appearance for each plt
newPlts <- newPlts[newPlts$Year!=2004,]                          #Remove 2004 (first year of data collection)
sz.cutoff <- 5                                                   #Sz cutoff, above which likely not recently a seedling 
newPlts.sm <- newPlts[newPlts$RosNew<sz.cutoff,]                 #Rm if larger than X ros (these likely missed & are not new)
newPlts.lrg <- newPlts[newPlts$RosNew>sz.cutoff,]                #Remove if smaller than X ros 
num.newPlts.sm <- newPlts.sm %>% group_by(Year) %>% summarise(NUM=n())   #Count num new plts per year
num.newPlts.lrg <- newPlts.lrg %>% group_by(Year) %>% summarise(NUM=n()) #Count num new plts per year

barplot(num.newPlts.sm$NUM, names=num.newPlts.sm$Year, main="New small plants (<5ros) per year", ylab="New plants")
barplot(num.newPlts.lrg$NUM, names=num.newPlts.lrg$Year, main="New large plants (>5ros) per year", ylab="New plants") 


## PLOT SMALL PLANTS IN EACH YEAR *****
erbr.sm <- erbr[erbr$RosNew<5 & !is.na(erbr$RosNew),]
num.erbr.sm <- erbr.sm %>% group_by(Year) %>% summarise(NUM=n())   #Count num small plts per year
barplot(num.erbr.sm$NUM, names=num.erbr.sm$Year, main="Small plants (<5ros) per year", ylab="Small plants")

```




## NUMBER OF PREDICTED INDIVS FOR EACH TRANSECT
```{r}

## Load matrices
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_P1k_210511")

names.mx <- unique(names(mx.list))
names.trans <- unique(sub('.*\\_', '', names.mx))     #Keep text following '_'
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
names.param <- unique(sub('\\-.*', '', names.mx))     #Keep text preceding '-'

## Subset mx.list for desired years if needed -----------------------------
names.yr <- names.yr[3:20] #Keep just 2004-2021
mx.studyYrs <- mx.list[grep(paste(names.yr, collapse="|"), names(mx.list))]
## ------------------------------------------------------------------------

sim.yrs <- length(names.yr)


## OBTAIN STABLE STAGE STRUCTURE
## Modified from Dan's code 
N.startNum <- 148                         #Set as observed population size in 2004

mx.num <- length(mx.studyYrs)
mx.mean <- Reduce('+', mx.studyYrs)/mx.num    #Calculate mean of all matrices 
mnees=eigenall(mx.mean)
N.vecStart=N.startNum * mnees$stable.stage


pop.sz.all <- NULL


## Start parameter loop (for a given parameter draw, loop over data yrs)
for (pp in 1:length(names.param)) {
  ## Subset by parameter
  mx.param <- mx.studyYrs[grep(paste(names.param[pp],"-",sep=""), names(mx.studyYrs))]
  pop.sz <- as.data.frame(matrix(NA, nrow=length(names.trans), ncol=sim.yrs))
  

  ## Transect loop
  for (tt in 1:length(names.trans)) {
  mx.paramTrans <- mx.param[grep(names.trans[tt], names(mx.param))]


  ## AG modified from Dan
  N.vec <- N.vecStart                                    #Re-set to desired starting values

  ## Loop over years (from e.g. 2004-2021 to give estimates for 2005-2022)
    for (ii in 1:(sim.yrs)) {
  
      mx <- mx.paramTrans[ii]

    
      ## Multiply matrix by vec w num indivs per sz class (initially this is stbl stage struct)
      N.vec <- as.matrix(as_tibble(mx)) %*% N.vec        #This grows the population one year
      pop.sz[tt,ii]= sum(N.vec) 

    } # end of ii (years) loop
  } # end of tt (transect) loop

   pop.sz.all <- rbind(pop.sz.all, pop.sz)
} #End pp loop



## Add and rename columns
pop.sz.all$TRANS <- rep(names.trans)
pop.sz.all$PARAM <- rep(names.param, each=length(names.trans))
pop.sz.all <- pop.sz.all[]
colnames(pop.sz.all) <- c(as.character(2005:2022),"TRANS","PARAM")
pop.sz.all$`2004` <- N.startNum
pop.sz.all <- pop.sz.all %>% relocate(`2004`, .before = `2005`)



## PLOT PREDICTED FOR EACH TRANSECT
cols <- viridis(12)

par(mar=c(4.5,5,3,2))
plot(NA,NA, ylim=c(0,180), xlim=c(2004,2022), ylab="Number of individuals", cex.lab=1.6,
     cex.axis=1.3,xlab="Year")
for (tt in 1:length(names.trans)) {
  pop.sz.trns <- subset(pop.sz.all, TRANS==names.trans[tt])
  pop.sz.means.trns <- colMeans(pop.sz.trns[1:((length(names.yr)+1))])
  pop.sz.sds.trns <- matrixStats::colSds(as.matrix(pop.sz.trns[1:((length(names.yr)+1))]))
  
  lines(2004:2022, pop.sz.means.trns[1:length(pop.sz.means.trns)],
        col=cols[tt],lwd=1.5)
  arrows(2004:2022,pop.sz.means.trns[1:length(pop.sz.means.trns)], 
         2004:2022,pop.sz.means.trns[1:length(pop.sz.means.trns)]+
         pop.sz.sds.trns[1:length(pop.sz.means.trns)], 
         length=0.05, angle=90, col=cols[tt])
  arrows(2004:2022,pop.sz.means.trns[1:length(pop.sz.means.trns)], 
         2004:2022,pop.sz.means.trns[1:length(pop.sz.means.trns)]-
         pop.sz.sds.trns[1:length(pop.sz.means.trns)], 
         length=0.05, angle=90, col=cols[tt])
}

legend("bottomleft", names.trans, col=cols,lwd=2,cex=1,
       horiz=FALSE, bty="y",seg.len=1)

```




## UPDATED PLOT OF ESTIMATED POP SZ PER YEAR COMPARED TO OBSERVED
```{r}

## Calculate observed: Subset by year and count the number of living plants 
num.indivs <- erbr %>% group_by(Year) %>% summarise(numIndivs=sum(surv, na.rm=TRUE)) 
num.indivs$Year <- as.numeric(as.character(num.indivs$Year))
num.indivs <- num.indivs[which(num.indivs$numIndivs!=0),]
## Do counts excluding anything less than given size (e.g. 5 ros)
erbr.lrg <- erbr[erbr$RosNew>5,]
#length(erbr.yr$RosNew[erbr.yr$RosNew<2 & !is.na(erbr.yr$RosNew)])
num.indivs <- erbr.lrg %>% group_by(Year) %>% summarise(numIndivs=sum(surv, na.rm=TRUE)) 
num.indivs$Year <- as.numeric(as.character(num.indivs$Year))
num.indivs <- num.indivs[which(num.indivs$numIndivs!=0),]


## To obtain predicted pop size, estimate for each transect separately and then sum together. 
## Load matrices
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_P1k_20230503")
mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to13_P1k_20231026")
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to13even_P1k_20231026")
#mx.list <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrIPMmxList_noYRE_4to13odd_P1k_20231026")
#mx.list <- mx.list4to13

names.mx <- unique(names(mx.list))
names.trans <- unique(sub('.*\\_', '', names.mx))     #Keep text following '_'
#names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
#names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
names.param <- unique(sub('\\-.*', '', names.mx))     #Keep text preceding '-'


## Subset to keep only years from study period (e.g. 2004-2021) (OPTIONAL?)
startYr <- "2007"
match.yr <- as.character(c(startYr:2021))
mx.list <- mx.list[grep(paste(match.yr,collapse="|"), names(mx.list))]
names.mx <- unique(names(mx.list))
names.yr <- unique(sub('\\_.*', '', names.mx))        #Keep text preceding '_'
names.yr <- unique(sub('.*\\-', '', names.yr))        #Keep text following '-'
mx.studyYrs <- mx.list[grep(paste(names.yr, collapse="|"), names(mx.list))]
sim.yrs <- length(names.yr)
## ---------------------------------------------------------------------------




## Determine the observed number of plts of each size for each transect at the start of the study (2004 or maybe choose 2007)
pltsStart <- erbr[erbr$Year==startYr,]  
hist(pltsStart$RosNew)
unique(pltsStart$RosNew)
table(pltsStart$RosNew)

minsize <- 1
maxsize <- max(erbr$RosNew, na.rm=TRUE)
bin.num <- 50
vec.bin <- c(minsize, minsize+1:bin.num*(maxsize-minsize)*(1/bin.num)) ##**SHOULD WE SET HIGHER MIN SZ? ##
vec.bin <- c(0,vec.bin) #Add additional zero-1 bin for seedlings
length(vec.bin)

## Loop over vec bin and assign bin bounds, then count num plts that fall into each size bin for each transect
binCounts = as.data.frame(matrix(NA, nrow=length(vec.bin)-1, ncol=length(names.trans)))
colnames(binCounts) <- names.trans

for (cc in 1:length(names.trans)) {
  pltsStartTran <- pltsStart[pltsStart$TransectNew==names.trans[cc],]
  
  for (bb in 1:length(vec.bin)-1) {
    bounds <- c(vec.bin[bb], vec.bin[bb+1])
    subsetSzs <- pltsStartTran[which(pltsStartTran$RosNew > bounds[1] & pltsStartTran$RosNew <= bounds[2]),]
      binCounts[bb,cc] <- nrow(subsetSzs)
  }
}
totsTrans <- colSums(binCounts)
#startNum <- sum(totsTrans)
startNum <- num.indivs$numIndivs[num.indivs$Year==startYr]


## Start parameter loop (for a given parameter draw, loop over transects and data yrs)
pop.sz.all <- NULL  #Variable for saving output

for (pp in 1:length(names.param)) {
  ## Subset by parameter
  mx.param <- mx.studyYrs[grep(paste(names.param[pp],"-",sep=""), names(mx.studyYrs))]
  pop.sz <- as.data.frame(matrix(NA, nrow=length(names.trans), ncol=sim.yrs))
  
  ## Transect loop
  for (tt in 1:length(names.trans)) {
  mx.paramTrans <- mx.param[grep(names.trans[tt], names(mx.param))]

  ## AG modified from Dan
  N.vec <- binCounts[,tt]                                    #Re-set to desired starting values
  N.vec[1:2] <- 0                                            #**TRY EXCLUDING ANYTHING >=4 rosettes; set to zero


  ## Loop over years (from e.g. 2004-2021 to give estimates for 2005-2022)
    for (ii in 1:(sim.yrs)) {
  
      mx <- mx.paramTrans[ii]
    
      ## Multiply matrix by vec w num indivs per sz class (initially this was observed nums)
      N.vec <- as.matrix(as_tibble(mx)) %*% N.vec        #This grows the population one year
      N.vec[1:2] <- 0                                    #**TRY EXCLUDING ANYTHING >=4 rosettes; set to zero
      pop.sz[tt,ii]= sum(N.vec) 

    } # end of ii (years) loop
  } # end of tt (transect) loop

   pop.sz.all <- rbind(pop.sz.all, pop.sz)
} #End pp (param) loop


## Add and rename columns
pop.sz.all$TRANS <- rep(names.trans)
pop.sz.all$PARAM <- rep(names.param, each=length(names.trans))
#pop.sz.all <- pop.sz.all[]
colnames(pop.sz.all) <- c(as.character((as.numeric(startYr)+1):2022),"TRANS","PARAM")


## Calculate mean and var of predicted pop size
## Sum across transects and average across params 
pop.sz.params <- pop.sz.all %>% group_by(PARAM) %>% summarise(across(c(1:(length(pop.sz.all)-2)), sum, na.rm=TRUE))
pop.sz.transMn <- colMeans(pop.sz.params[,2:length(pop.sz.params)])
pop.sz.transSd <- matrixStats::colSds(as.matrix(pop.sz.params[,2:length(pop.sz.params)]))
pop.sz.transMn <- c(startNum, pop.sz.transMn)  #Add observed num at start
pop.sz.transSd <- c(0, pop.sz.transSd)



## SAVE FOR PLOTTING BELOW BY RENAMING BASED ON DATA SET 
pop.sz.means.full <- pop.sz.transMn
pop.sz.sds.full <- pop.sz.transSd
names.yr.full <- names.yr

pop.sz.means.consec <- pop.sz.transMn
pop.sz.sds.consec <- pop.sz.transSd
names.yr.consec <- names.yr

#pop.sz.means.evn <- pop.sz.transMn
#pop.sz.sds.evn <- pop.sz.transSd
#names.yr.evn <- names.yr

#pop.sz.means.odd <- pop.sz.transMn
#pop.sz.sds.odd <- pop.sz.transSd
#names.yr.odd <- names.yr
#pop.sz.ciHi.odd <- pop.sz.ciHi
#pop.sz.ciLo.odd <- pop.sz.ciLo


## Save output 
#setwd("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship")
dateToday <- Sys.Date()        #Get date to be added to file name
dateToday <- str_replace_all(dateToday, "-", "") 
saveRDS(pop.sz.means.full, file=paste("erbrPopSzMnFull07", dateToday, sep="_"))
saveRDS(pop.sz.sds.full, file=paste("erbrPopSzSdFull07", dateToday, sep="_"))
saveRDS(pop.sz.means.consec, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMn4to13-07", dateToday, sep="_"))
saveRDS(pop.sz.sds.consec, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSd4to13-07", dateToday, sep="_"))
saveRDS(pop.sz.means.evn, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMn4to13evn-07", dateToday, sep="_"))
saveRDS(pop.sz.sds.evn, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSd4to13evn-07", dateToday, sep="_"))
saveRDS(pop.sz.means.odd, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMn4to13odd-07", dateToday, sep="_"))
saveRDS(pop.sz.sds.odd, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSd4to13odd-07", dateToday, sep="_"))
## So can load and re-calc var (e.g. as quantiles instead of SD) later and re-plot
saveRDS(pop.sz.all, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzAllFull07", dateToday, sep="_"))
saveRDS(pop.sz.all, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzAll4to13odd-07", dateToday, sep="_"))
saveRDS(pop.sz.all, 
        file=paste("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzAll4to13-07", dateToday, sep="_"))



## PLOT PREDICTED VS OBSERVED POP SZ
## Read in saved files
pop.sz.means.full <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMnFull07_20231229") 
pop.sz.sds.full <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSdFull07_20231229")  
pop.sz.means.consec <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMn4to13-07_20231230") 
pop.sz.sds.consec <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSd4to13-07_20231230")  
#pop.sz.means.evn <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMn4to13evn-07_20231230") 
#pop.sz.sds.evn <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSd4to13evn-07_20231230")  
#pop.sz.means.odd <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzMn4to13odd-07_20231230") 
#pop.sz.sds.odd <- readRDS("C:/Users/april/Dropbox/CU_Boulder_PhD/DBG_Internship/erbrPopSzSd4to13odd-07_20231230")  


## Plot mean and SD (or CIs etc)
#plot(as.numeric(as.character(num.indivs$Year)), num.indivs$numIndivs, type='l', ylim=c(0,250), 
#     ylab="Number of individuals",xlab="Year", cex.lab=1.45, cex.axis=1.3)
#points(as.numeric(as.character(num.indivs$Year)), num.indivs$numIndivs, pch=16, cex=0.75)

#lines(2004:2022,pop.sz.means.full[1:19], #pop.sz.means.full[1:length(pop.sz.means.full)]
#      col="grey60",lwd=1.9)  
#arrows(2004:2022,pop.sz.means.full[1:19], 
#       2004:2022,pop.sz.means.full[1:19]+pop.sz.sds.full[1:19], 
#       length=0.05, angle=90, col="grey60", lwd=0.6)
#arrows(2004:2022,pop.sz.means.full[1:19], 
#       2004:2022,pop.sz.means.full[1:19]-pop.sz.sds.full[1:19], 
#       length=0.05, angle=90, col="grey60", lwd=0.6)

#legend("topright", c("Observed", "Full dataset"), col=c("black","grey60"),lwd=2.95,cex=1,
#       horiz=FALSE, bty="y",seg.len=1, xpd=TRUE)



## Plot from 2007 onward
pdf('20240126_ErBr_fig4.pdf', width=6.5, height=4.5, pointsize=8)
plot(as.numeric(as.character(num.indivs$Year[4:14])), num.indivs$numIndivs[4:14], type='l', ylim=c(30,150), 
     ylab="Number of individuals",xlab="Year", cex.lab=1.2, cex.axis=1.1, lwd=2)

lines(2007:2022, pop.sz.means.consec[1:length(pop.sz.means.consec)],
      col="#fc8d59",lwd=2)
arrows((2007:2022),pop.sz.means.consec[1:length(pop.sz.means.consec)], 
       (2007:2022),pop.sz.means.consec[1:length(pop.sz.means.consec)]+pop.sz.sds.consec[1:length(pop.sz.means.consec)], 
       length=0, angle=90, col="#fc8d59", lwd=0.6)
arrows((2007:2022),pop.sz.means.consec[1:length(pop.sz.means.consec)], 
       (2007:2022),pop.sz.means.consec[1:length(pop.sz.means.consec)]-pop.sz.sds.consec[1:length(pop.sz.means.consec)], 
       length=0, angle=90, col="#fc8d59", lwd=0.6)

points(as.numeric(as.character(num.indivs$Year)), num.indivs$numIndivs, pch=16, cex=0.85)

lines(2007:2022,pop.sz.means.full[1:16], #pop.sz.means.full[1:length(pop.sz.means.full)]
      col="#d73027",lwd=2)  
arrows(2007:2022,pop.sz.means.full[1:16], 
       2007:2022,pop.sz.means.full[1:16]+pop.sz.sds.full[1:16], 
       length=0, angle=90, col="#d73027", lwd=0.6)
arrows(2007:2022,pop.sz.means.full[1:16], 
       2007:2022,pop.sz.means.full[1:16]-pop.sz.sds.full[1:16], 
       length=0, angle=90, col="#d73027", lwd=0.6)

legend("bottomleft", c("Observed", "Full dataset", "2004-2013"), col=c("black","#d73027","#fc8d59"),lwd=2,cex=1.2,
       horiz=FALSE, bty="y",seg.len=1, xpd=TRUE)

dev.off()

#lines(2007:2022, pop.sz.means.odd[1:length(pop.sz.means.odd)],
#      col="#4575b4",lwd=1.9)
#arrows((2007:2022)-0.1,pop.sz.means.odd[1:length(pop.sz.means.odd)], 
#       (2007:2022)-0.1,pop.sz.means.odd[1:length(pop.sz.means.odd)]
#       +pop.sz.sds.odd[1:length(pop.sz.means.odd)],length=0.05, angle=90, col="#4575b4", lwd=0.6)
#arrows((2007:2022)-0.1,pop.sz.means.odd[1:length(pop.sz.means.odd)], 
#       (2007:2022)-0.1,pop.sz.means.odd[1:length(pop.sz.means.odd)]
#       -pop.sz.sds.odd[1:length(pop.sz.means.odd)],length=0.05, angle=90, col="#4575b4", lwd=0.6)

#lines(2007:2022, pop.sz.means.evn[1:length(pop.sz.means.evn)],
#      col="#91bfdb",lwd=1.9)
#arrows((2007:2022)+0.1,pop.sz.means.evn[1:length(pop.sz.means.evn)], 
#       (2007:2022)+0.1,pop.sz.means.evn[1:length(pop.sz.means.evn)]+pop.sz.sds.evn[1:length(pop.sz.means.evn)], 
#       length=0.05, angle=90, col="#91bfdb", lwd=0.6)
#arrows((2007:2022)+0.1,pop.sz.means.evn[1:length(pop.sz.means.evn)], 
#       (2007:2022)+0.1,pop.sz.means.evn[1:length(pop.sz.means.evn)]-pop.sz.sds.evn[1:length(pop.sz.means.evn)], 
#       length=0.05, angle=90, col="#91bfdb", lwd=0.6)

#,"#91bfdb","#4575b4"
#, "2004-2013 even yrs only", "2004-2013 odd yrs only"

```





## PLOT THE DISTRIBUTION OF OBSERVED PLANT SIZES
```{r}

head(erbr)
#using RosNew, look at distribution of sizes across all years
hist(erbr$RosNew, breaks=100, xlim=c(0,150), xlab="Plant size (rosette number)", 
     ylab="Number of individuals", main="Distribution of plant sizes across all years")

#also look at distribution of sizes in one year at a time
erbr.yr <- erbr[erbr$Year==2007,]
hist(erbr.yr$RosNew, breaks=50, xlim=c(0,120))
erbr.yr <- erbr[erbr$Year==2010,]
hist(erbr.yr$RosNew, breaks=50, xlim=c(0,130))
erbr.yr <- erbr[erbr$Year==2018,]
hist(erbr.yr$RosNew, breaks=40, xlim=c(0,60))


## Look at number of plants less than given size in different years
erbr.yr <- erbr[erbr$Year==2007,]
length(erbr.yr$RosNew[erbr.yr$RosNew<2 & !is.na(erbr.yr$RosNew)])
erbr.yr <- erbr[erbr$Year==2008,]
length(erbr.yr$RosNew[erbr.yr$RosNew<2 & !is.na(erbr.yr$RosNew)])
erbr.yr <- erbr[erbr$Year==2009,]
length(erbr.yr$RosNew[erbr.yr$RosNew<2 & !is.na(erbr.yr$RosNew)])
erbr.yr <- erbr[erbr$Year==2010,]
length(erbr.yr$RosNew[erbr.yr$RosNew<2 & !is.na(erbr.yr$RosNew)])

```


